The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2016-02-02 DishBeamMeasurements/paper/DishBeamMeasurementsPaper_rev3.tex
incorporating comments

\shorttitle{The HERA Dish: Beam Measurements and Science Implications}
\shortauthors{Neben et al.}

@@

\author{Abraham R. Neben\altaffilmark{1},
Richard F. Bradley\altaffilmark{2,3},
Aaron Ewall-Wice\altaffilmark{1},
Nipanjana Patra\altaffilmark{4},
Nithyanandan Thyagarajan\altaffilmark{5},
James E. Aguirre\altaffilmark{6},
Zaki S. Ali\altaffilmark{4},
Judd Bowman\altaffilmark{5}
Carina Cheng\altaffilmark{4},
David R. DeBoer\altaffilmark{4}
Jacqueline N. Hewitt\altaffilmark{1},
Aaron Parsons\altaffilmark{4}}

@@

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral Hydrogen from the Epoch of Reionization. Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m) dishes with suspended dipole feeds with element collecting area of 100\,m$^2$. This is the first of a series of four papers characterizing the frequency and angular response of the HERA dish element with simulations and measurements. In this work we measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure collecting areas of 70--100\,m$^2$ and predict power spectrum SNRs. With optimistic foreground and analysis assumptions, HERA-127 should detect the EOR with a signal-to-noise of 25--30 with a single season of observations. Even pessimistically assuming that only previously demonstrated techniques are used, the significance of an EOR detection remains above $5\sigma$. Lastly we simulate foreground visibilities using numerical beam models and study the foreground distribution in Fourier space as a function of LST, and the uncertainties in these predictions due to beam modeling uncertainties.
\end{abstract}

@@

A new generation of low frequency radio telescopes is coming online with the goal of
 probing redshifted 21\,cm emission from the Cosmic Dawn. These observations will 
 complement indirect probes of the Dark Ages and Epoch of Reionization such as quasar 
 sightlines and the CMB optical depth which leave the reionization 
 history of the universe only loosely constrained. (See \citet{FurlanettoReview, miguelreview, PritchardLoebReview, aviBook, zaroubi} for reviews) Sensitivity and foreground removal are 
 the main challenges in 21\,cm observations, as the expected cosmological signal is 4--5 
 orders of magnitude fainter than Galactic and extragalactic foregrounds. Radio 
 interferometers such as the Murchison Widefield Arry (MWA) \citep{tingay13,mwascience}, the Precision Array for Probing the Epoch of Reionization (PAPER) \citep{ali2015}, the Giant Meterwave Radio Telescope (GMRT) 
 \citep{Paciga2011}, and the Low Frequency Array (LOFAR) \citep{lofar} are seeking a first detection of 
 cosmological 21\,cm emission in power spectrum measurements, where the smooth 
 frequency evolution of the foreground emission separates from the spectrally 
 unsmooth cosmological signal whose frequency dimension probes the a line of sight through the 
 inhomogenous reionizing universe.

The Hydrogen Epoch of Reionization Array (HERA) \citep[][, deBoer et al., submitted]{PoberNextGen} is drawing on lessons learned by the MWA and PAPER to reach the calibration and foreground isolation accuracy required to make a significant detection and characterization of the cosmological signal. HERA uses 14\,m diameter parabolic dishes arranged in a compact, hexagonal array to achieve coherent integration on the very low surface brightness 21\,cm signal. Redundant baselines also permit redundant calibration techniques which solve for the relative calibration between all antennas \citep{zheng14}, and use a sky model only to set the frequency dependent flux scale. In contrast, non-redundant arrays are pursuing fully sky model-based calibration schemes which risk frequency-dependent calibration systematics due to sky modeling inaccuracy \citep{braun2013}. HERA is pursuing a staged deployment of 19, 127, and finally 331 elements in progressively larger hex patterns, with scattered outriggers for imaging. A central lesson of first generation instruments is it is essential to characterize the instrument response to foreground emission lest instrument frequency dependence smear foreground power into cosmological signal modes.

@@

In an ideal achromatic instrument the foreground emission would be confined to the lowest few 
line of sight Fourier modes \citep[e.g.,][]{MoralesBowmanHewittFGsub}, however  the  
interferometer's frequency-dependent point spread function smears foreground power into a ``wedge'' shaped 
region in  $(k_\perp,k_\parallel)$ Fourier space \citep{Dattapowerspec,X13, PoberWedge,MoralesPSShapes, VedanthamWedge, nithya13, CathWedge, AdrianWedge1, AdrianWedge2}. This effect is straightforward to understand for a single baseline which measures the sky intensity weighted by the complex sky fringe 
$e^{i \vec{k}\cdot\vec{b}}$, where $\vec{k}=\vec{k}(\theta,\phi,f)$ is the wave vector of 
the incident radiation, $\vec{b}$ is the baseline vector in meters, and $f$ is the observation frequency. Thus sources at different positions relative to the baseline vector manifest different 
frequency structure despite their intrinsically smooth spectra, but are geometrically limited 
by the baseline length to a maximum frequency dependence of  $e^{2\pi i f b/c}$. This limits the foreground contamination 
to a wedge shaped region in Fourier space with $k_\parallel<a k_\perp$, where $k_\perp$ and $k_\parallel$ represent spatial modes 
perpendicular and parallel to the line of sight, and $a$ is a constant depending on the observational frequency and cosmology. The complement of the wedge is knows as the ``EOR window''.

It is convenient to phrase this description in terms 
of the delay in radiation arrival at the baseline's two antennas, $\tau$, where $\tau_
\text{max}=b/c$. Sources at low delay have little frequency 
structure, while those near $\tau=\tau_\text{max}$ acquire the maximum frequency 
structure given the baseline length.

The fact that sources acquire frequency dependence based on with their position on the sky 
tells us already that the primary beam strongly affects the aggregate frequency dependence 
of the foregrounds. The high delay regions of the sky lie near the horizon while low delay 
regions lie closer to zenith and also perpendicular to the baseline vector. \citet{nithya15} 
simulate the foreground contamination seen with a dipole beam, a phased array, and a Airy pattern,
and find that the latter suffers the least foreground leakage into
 $k_\parallel>0$ modes due to its narrow main lobe and minimal sidelobe 
levels. To be sure, all are subject to the same geometric limits on foreground frequency-
dependence limiting foreground bounding foreground emission within the wedge, but the emission from high delay is better suppressed using the 
Airy pattern leaving much of the wedge effectively empty.

@@

From the point of view of power spectrum estimation, so long as foreground emission is perfectly contained in the wedge it is irrelevant how much or 
little of it there is, but the finite bandwidth and imperfect bandpass calibration 
of real instruments smear power beyond the geometrical edge of the wedge into the EOR 
window. Sources at higher delay appear closest to the edge of the wedge, and thus are 
most at risk of leaking into the EOR window due to these effects. In fact, 
\citet{nithya15,nithya15b} observe in simulations and then in data that while naively we might expect minimal emission 
at the very edge of the wedge because typical near-horizon beam responses are so small, 
two effects can cause a relative brightening of emission at those maximal delays, creating a characteristic ``pitchfork'' shape. This horizon 
brightening is caused by the large solid angle subtended by the near-horizon regions of the 
sky, as well as the apparent shortening of baselines when viewed nearly on axis at these elevations. 
This second effect makes intermediate length baselines of tens to hundreds of meters sensitive to the very bright diffuse 
emission would not see from near zenith. Together, these effects can overcome the decline in beam sensitivity near the horizon. 
All these considerations highlight the antenna beam as a critical design parameter for 
21\,cm observatories

@@

This is the first in a series of four papers detailing the HERA element. In this work we study 
angular response of the dish and its implications for power spectrum measurements. The three companion 
papers present reflectometry measurements (Patra et al., submitted) and simulations (Ewall-Wice et al., 
submitted) of the dish frequency response, as well as detailed 
foreground simulations for HERA (Thyagarajan et al., submitted). A general description to the design of the 
HERA experiment from an engineering point of view is given by DeBoer et al. (submitted). In essence, we 
require a large collecting area for
 sensitivity and minimal sidelobes and horizon response without incurring the large cost per collecting 
area of very large dishes. A dish is preferred to a large phased array as it has fewer degrees 
of freedom and reduced potential of antenna-to-antenna variation (Neben et al 2015b, 
submitted). These factors naturally lead to a 14\,m diameter parabolic dish with a dipole 
feed suspended at prime focus. The 352 dishes are positioned in a compact, hexagonal 
array permitting redundant baseline calibration and coherent integration in $
\vec{k}$ space \citep{omniscope,ali2015}. 
%The smaller field of view of the HERA dish relative to that of the PAPER dipole or the MWA tile does not limit sensitivity in Fourier space as most leverage on high $k$ modes is from line of sight modes anyway \citep{PoberNextGen}.

@@

In this paper we first characterize the angular response of a prototype HERA dish at the National Radio 
Astronomy Observatory--Green Bank. We use the beam mapping system of \citet{neben15} to 
measure the 137\,MHz beam pattern using the ORBCOMM satellite constellation. We obtain beam 
measurements out to zenith angles of $\sim60^\circ$ where the beam response is -35\,dB relative to zenith, and compare with different numerical models. We characterize the dish beam at various feed heights to map out the focus and study beam errors due to feed misalignment. We compute the collecting areas and implied EOR power spectrum sensitivities of our measured beams. After verifying our numerical models, we consider the science implications of these 
beam patterns by foreground delay spectra at different baseline lengths and observing conditions to study when the horizon brightening effect is strongest, and thus, when foreground are must at risk of leaking into the EOR window.

In detail, we discuss the electromagnetic design and modeling of the dish in Section 2. We present the 
experimental setup of the beam mapping experiments and discuss their systematics, then 
review the ORBCOMM beam measurement system, in Section 3. We present our power pattern 
measurements in Section 4, and study the science implications of these beam measurements for foreground power spectra in Section 5, then conclude with discussion in Section 6.

@@

We review here the design of the HERA dish, and refer to DeBoer et al. (submitted) for details. 
The 14\,m HERA dish design is a departure from the large N--small D approach used by 21\,cm 
observatories like PAPER and the MWA. Both observatories are actively pursuing power 
spectrum analyses using several year data runs, but the sheer data volume makes 
characterization and removal of systematics, as well as repeated or complimentary analyses, 
challenging. A larger element was chosen for HERA primarily to reduce correlation and data processing costs.

The HERA element is a 14\,m diameter parabolic dish ($f/D\approx0.32$) with a dual-polarized dipole feed (Fig. \ref{fig:feedphoto}) suspended at prime focus. The dish surface is formed by sheets of wire attached to PVC tubes attached to supports at the edge of the dish and to a hub at the base. The feed is suspended with cables attached to three telephone poles around the dish. For feed maintenance purposes, a door is engineered into the dish surface on one mesh panel. The dipole stands 17\,in below a 78\,in diameter circular mesh back plane surrounded by a 30\,in deep cylinder. The cylinder is offset 0.5\,in from the back plane. The feed is designed to make the beams for both polarizations more similar to each other, and also to taper the dipole beam towards the edge of the dish to mitigate dish-dish coupling.  The nominal dish focus is at 4.5\,m from the dish surface, though numerical dish models including the feed predict the focus is closer to 5.3\,m.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Prototype HERA feed set up outside the dish for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17\,in from a 78\,in diameter back plane, surrounded by a 30\,in deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

One cost of using fewer large antenna elements is a smaller field of view, but this is a 
small effect for 21\,cm power spectrum analyses as our leverage on $k$ modes comes primarily from $k_\parallel$ modes along the line of sight (in the frequency dimension). More significant is frequency structure introduced by time domain reflections between the suspended feed detailed by Ewall-Wice et al. (submitted) with simulations and Patra et al. (submitted) with zenith reflectometry measurements.

@@

We numerically model the HERA dish in both HFSS and in CST, making slightly different assumptions, in order to study the dependence of the beam pattern on modeling imperfections. In particular, we suspect that the near horizon beam response, which sets the level of horizon brightening, is quite sensitive to modeling assumptions. Modeling is also useful to probe the best focus of the dish--feed system. The nominal focus is at 4.5\,m given that the dish has $f/D=0.32$, though the feed back plane and skirt are expected to move the focus slightly.

The HFSS model uses a 14\,m paraboloid with $f/D=0.32$ as the dish, with a 1\,m hole at vertex. In reality the 1\,m diameter at the vertex is filled with concrete and supports the PVC pipes which suspend the mesh panels, but we effectively assume here it is perfectly absorbant.  The feed is modeled as a PAPER dipole suspended 17\,in below a 78\,in diameter circular backplane inside a 30\,in deep skirt hanging 0.5\,in below the backplane. All these surfaces are modeled as solid aluminum, as are the dipole ``sleeves'', circular disks just above and below the dipole plane designed to broaden its frequency response. The dipoles themselves is modeled as a copper pipe of diameter XX, we neglect the dielectric PVC in which the copper pipe is actually enclosed. In the simulations, we excite one of the dipoles using a modal port and measure the total gain response in each direction to 137.5\,MHz radiation.

In the CST model we again model the dish as a 14\,m diameter solid aluminum paraboloid with $f/D=0.32$ but now neglect the 1\,m hole at zenith, effectively assuming the concrete and earth ground behind it are perfectly reflective. The actual material properties are somewhere in between, though difficult to predict theoretically. The feed model has the same dimensions as in the HFSS model. The simulations are done by exciting one of the dipoles with band-limited noise (100--200\,MHz). The beam pattern is then obtained at a given frequency using a farfield monitor after the excitation pulse energy within the structure decays to below -80\,dB.

@@

\subsection{ORBCOMM Beam Mapping System Review}

We briefly review the beam mapping system detailed by \citet{neben15}, then discuss 
application of system for HERA dish measurements. The system 
takes advantage of the 137\,MHz communications satellites operated by ORBCOMM Inc. 
as bright point sources which, by virtue of their number ($\sim30$), short orbital periods 
($\sim90$ minutes), and orbital precession cover $\sim$65\% of the visible sky in just a few 
days. The coverage from the Green Bank site is limited by the fact that the satellites' orbital inclinations are all less 
than $45^\circ$.

@@

In detail, we measure the dual-polarization RMS power received by each antenna in 512 2\,kHz 
channels across the 137--138\,MHz band. Each band power is averaged over $\sim0.2$
\,sec. There are 0--3 satellites above the horizon at any given time transmitting on different 
$\sim15$\,kHz wide sub-bands in 137--138\,MHz. By observing at many different 
frequencies, we probe the beam response in all these directions simultaneously. We 
compute the satellite positions using the orbital elements published by Celestrak
\footnote{http://www.celestrak.com/NORAD/elements/orbcomm.txt} and the orbital 
integrator \texttt{predict}\footnote{http://www.qsl.net/kd2bd/predict.html}. However, the 
satellite frequencies vary occasionally to avoid interference within the constellation. 
\citet{zheng14} use interferometric phases to identify and exclude times when multiple 
satellites are in view. As our data acquisition system makes only total power 
measurements, we instead use an ORBCOMM interface box (typically supplied to 
commercial users of the network) to sync with passing satellites and record their identifier 
and transmission frequency during each pass.

In this way, beam measurements are built up along satellite tracks over the course of 
several days of integration, yielding typically 200--300 satellite pass. Each pass is 
processed separately to identify and exclude times of low signal-to-background when the 
satellite is low in the sky or in the off state of of a pulsing sequence. At those times, then 
satellite flux no longer dominates over that of the diffuse Galactic background, and a 
power measurement no longer probes the response in only the satellite direction. The beam 
measurements are then gridded in horizontal coordinates in HEALPix \citep{healpix} with a resolution of 
$1.8^\circ$ (nside=32). As a last quality control step to reject errant beam measurements 
due to RFI, for instance, we keep only the central 90\% of $\sim50$ measured beam 
values in each HEALPix cell.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{IMG_0353.JPG}
\caption{The dish with its suspended feed is seen in the back, 50\,m north of one of the reference antennas used in the null experiment to study systematics. The experiment is conducted in Galford Meadow at NRAO--Green Bank.}
\label{fig:greenbankdishphoto}
\end{figure}

We use a simple dual-polarization dipole as our reference antenna. The dipole is constructed out of copper tubing covered by PVC for protection, mounted above a 2\,m $\times$ 2\,m ground plane. See \citet{neben15} for details. During the dish measurements the dipole is positioned 100\,m due south of the dish, though we experiment with other locations at first in our to characterize the environmental systematics of these measurements, as detailed in the next section. Figure \ref{fig:greenbankdishphoto} shows the dish with suspended feed 50\,m north of one of the reference antennas.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_maps.pdf}
\caption{Measured dish power patterns at various feed heights for the EW (left panel) and NS (right panel) instrumental polarizations. The sidelobes shrink and the main lobe narrows as the feed is raised, confirming that the focus of the dish/feed system is higher than the 4.5\,m ideal focus of the parabolic dish.}
\label{fig:measuredbeammaps}
\end{figure*}

>>>>

\shorttitle{The HERA Dish I: Beam Measurements and Science Implications}
\shortauthors{Neben et al.}

@@

\author{Abraham R. Neben\altaffilmark{1},
Richard F. Bradley\altaffilmark{2,3},
Jacqueline N. Hewitt\altaffilmark{1},
David R. DeBoer\altaffilmark{4},
Aaron Parsons\altaffilmark{4},
Aaron Ewall-Wice\altaffilmark{1},
Nipanjana Patra\altaffilmark{4},
Nithyanandan Thyagarajan\altaffilmark{5},
James E. Aguirre\altaffilmark{6},
Zaki S. Ali\altaffilmark{4},
Judd Bowman\altaffilmark{5},
Carina Cheng\altaffilmark{4}
Saul A. Kohn\altaffilmark{6},
Roger Dickenson\altaffilmark{3},
Phillip Doolittle\altaffilmark{3},
Dennis Egan\altaffilmark{3},
Mike Hedrick\altaffilmark{3},
Patrick Schaffner\altaffilmark{3},
John Shelton\altaffilmark{3},
H.A. Taylor\altaffilmark{3},
Rusty Taylor\altaffilmark{3},
Butch Wirt\altaffilmark{3}}

@@

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral hydrogen from the Epoch of Reionization. Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m) dishes with suspended dipole feeds with element collecting area of 100\,m$^2$. Not only does the dish determine overall sensitivity, it determines the aggregate frequency structure of foregrounds in the interferometer. This is the first of a series of four papers characterizing the frequency and angular response of the with simulations and measurements. We focus in this paper on the angular response (i.e., power pattern), which sets the relative weighting between sky regions of high and low delay, and thus, apparent source frequency structure. We measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure a collecting area of 97\,m$^2$ in the optimal dish/feed configuration, implying HERA-331 should detect the EOR with a signal-to-noise of 74 with a single season of observations, making optimistic foreground and analysis assumptions. Even with more pessimistic assumptions, using only previously demonstrated techniques, a HERA should achieve a signal-to-noise of 20. Lastly we simulate foreground visibilities using numerical beam models and study the foreground distribution in Fourier space as a function of local sidereal time, and the uncertainties in these predictions due to beam modeling uncertainties.
\end{abstract}

@@

A new generation of low frequency radio telescopes is coming online with the goal of
 probing redshifted 21\,cm emission from the Cosmic Dawn. These observations will 
 complement indirect probes of the Dark Ages and Epoch of Reionization such as quasar 
 sightlines and the CMB optical depth which leave the reionization 
 history of the universe only loosely constrained. (See \citet{FurlanettoReview, miguelreview, PritchardLoebReview, aviBook, zaroubi} for reviews) Sensitivity and foreground removal are 
 the main challenges in 21\,cm observations, as the expected cosmological signal is 4--5 
 orders of magnitude fainter in brightness temperature than Galactic and extragalactic foregrounds. Radio 
 interferometers such as the Murchison Widefield Arry (MWA) \citep{tingay13,mwascience}, the Precision Array for Probing the Epoch of Reionization (PAPER) \citep{ali2015}, the Giant Meterwave Radio Telescope (GMRT) 
 \citep{Paciga2011}, and the Low Frequency Array (LOFAR) \citep{lofar} are seeking a first detection of 
 cosmological 21\,cm emission in power spectrum measurements, where the smooth 
 frequency evolution of the foreground emission separates from the spectrally 
 rough cosmological signal whose frequency dimension probes a line of sight through the 
 inhomogenous reionizing universe.

The Hydrogen Epoch of Reionization Array (HERA) \citep[][deBoer et al. (in prep)]{PoberNextGen} is drawing on lessons learned by the MWA and PAPER to reach the calibration and foreground isolation accuracy required to make a significant detection and characterization of the cosmological signal. HERA uses 14\,m diameter parabolic dishes arranged in a compact, hexagonal array to achieve coherent integration on the very low surface brightness 21\,cm signal. Redundant baselines also permit redundant calibration techniques which solve for the relative calibration between all antennas \citep{liu2010,zheng14}. In contrast, non-redundant arrays are pursuing fully sky model-based calibration schemes which risk frequency-dependent calibration systematics due to sky modeling inaccuracy \citep{braun2013}. A central lesson from first generation instruments is that it is essential to characterize the instrument response to foreground emission lest instrument frequency dependence smear foreground power into cosmological signal modes.

@@

In an ideal achromatic instrument the foreground emission would be confined to the lowest few 
line of sight Fourier modes \citep[e.g.,][]{MoralesBowmanHewittFGsub}, however  the  
interferometer's frequency-dependent point spread function smears foreground power into a ``wedge'' shaped 
region in  $(k_\perp,k_\parallel)$ Fourier space \citep{Dattapowerspec,X13, PoberWedge,MoralesPSShapes, VedanthamWedge, nithya13, CathWedge, AdrianWedge1, AdrianWedge2,parsons12b}. This effect is straightforward to understand for a single baseline which measures the sky intensity weighted by the complex sky fringe 
$e^{i \vec{k}\cdot\vec{b}}$, where $\vec{k}(f)$ is the wave vector of incident radiation pointing along its direction of travel, $\vec{b}$ is the baseline vector in meters, and $f$ is the observation frequency. Thus sources at different positions relative to the baseline vector manifest different 
frequency structure despite their intrinsically smooth spectra, but are geometrically limited 
by the baseline length to a maximum frequency dependence of  $e^{2\pi i f b/c}$. This limits the foreground contamination 
to a wedge shaped region in Fourier space with $k_\parallel<a k_\perp$, where $k_\perp$ and $k_\parallel$ represent spatial modes 
perpendicular and parallel to the line of sight, and $a$ is a constant depending on the observational frequency and cosmology. The complement of the wedge is knows as the ``EOR window''.

It is convenient to phrase this description in terms 
of the delay in radiation arrival at the baseline's two antennas, $\tau$, where $\tau_
\text{max}=b/c$. Sources at low delay have little frequency 
structure, while those near $\tau=\tau_\text{max}$ acquire the maximum frequency 
structure given the baseline length. The high delay regions of the sky lie near the horizon while low delay 
regions lie closer to zenith and also perpendicular to the baseline vector.

So because sources acquire frequency dependence based on their position on the sky, and the primary beam weights different regions of the sky differently, we see that the primary beam (i.e., the antenna angular response) strongly affects the aggregate frequency dependence 
of the foregrounds. \citet{nithya15} simulate the foreground contamination seen with a dipole beam, a phased array, and a Airy pattern,
and find that the latter suffers the least foreground leakage into
 $k_\parallel>0$ modes due to its narrow main lobe and minimal sidelobe 
levels. To be sure, all are subject to the same geometric limits on foreground frequency-
dependence limiting foreground bounding foreground emission within the wedge, but the emission from high delay is better suppressed using the 
Airy pattern leaving much of the wedge effectively empty.

@@

For foreground avoidance-based power spectrum estimation, so long as foreground emission is perfectly contained in the wedge it is irrelevant how much or 
little of it there is, but real world effects smear power beyond the geometrical edge of the wedge into the EOR 
window. Finite bandwidth, imperfect bandpass calibration, and faraday rotation of polarized sources can all imprint slight frequency structure on otherwise spectrally smooth sources, and those appearing closest to the edge of the wedge are 
most at risk of leaking into the EOR window. In fact, 
\citet{nithya15,nithya15b} observe in simulations and then in data that while naively we might expect minimal emission 
at the very edge of the wedge because typical near-horizon beam responses are so small, 
two effects can cause a relative brightening of emission at those maximal delays, creating a characteristic ``pitchfork'' shape. This horizon 
brightening is caused by the large solid angle subtended by the near-horizon regions of the 
sky, as well as the apparent shortening of baselines when viewed nearly on axis at these elevations. 
This second effect makes intermediate length baselines of tens to hundreds of meters sensitive to the very bright diffuse 
emission they would not see from near zenith. Together, these effects can overcome the decline in beam sensitivity near the horizon. 
All these considerations highlight the antenna beam as a critical design parameter for 
21\,cm observatories.

@@

This is the first in a series of four papers detailing the HERA element. In this work we study 
angular response of the dish and its implications for power spectrum measurements. The three companion 
papers present reflectometry measurements (Patra et al., submitted) and simulations (Ewall-Wice et al., 
submitted) of the dish frequency response, as well as detailed 
foreground simulations for HERA (Thyagarajan et al., submitted). A general description to the design of the 
HERA experiment is given by DeBoer et al. (in prep). In essence, we 
require a large collecting area for
 sensitivity and minimal sidelobes and horizon response without incurring the large cost per collecting 
area of very large dishes. A dish is preferred to a large phased array as it has fewer degrees 
of freedom and reduced potential of antenna-to-antenna variation (Neben et al 2015b, 
submitted). The 331 dishes are positioned in a compact, hexagonal 
array permitting redundant baseline calibration and coherent integration in $
\vec{k}$ space \citep{omniscope,parsons12a}. 
%The smaller field of view of the HERA dish relative to that of the PAPER dipole or the MWA tile does not limit sensitivity in Fourier space as most leverage on high $k$ modes is from line of sight modes anyway \citep{PoberNextGen}.

@@

In this paper we first characterize the angular response of a prototype HERA dish at the National Radio 
Astronomy Observatory--Green Bank. We use the beam mapping system of \citet{neben15} to 
measure the 137\,MHz beam pattern using the ORBCOMM satellite constellation. We obtain beam 
measurements out to zenith angles of $\sim60^\circ$ where the beam response is -35\,dB relative to zenith, and compare with different numerical electromagnetic models. We characterize the dish beam at various feed heights to map out the focus and study beam errors due to feed misalignment. We compute the collecting areas and implied EOR power spectrum sensitivities of our measured beams. After verifying our models, we consider the science implications of these 
beam patterns by foreground delay spectra at different baseline lengths and observing conditions to study when the horizon brightening effect is strongest, and thus, when foreground are most at risk of leaking into the EOR window.

We discuss the electromagnetic design and modeling of the dish in Section 2. We present the 
experimental setup of the beam mapping experiments and discuss their systematics, then 
review the ORBCOMM beam measurement system, in Section 3. We present our power pattern 
measurements in Section 4, and study the science implications of these beam measurements for foreground power spectra in Section 5, then conclude with discussion in Section 6.

@@

We review here the design of the HERA dish, and refer to DeBoer et al. (in prep) for details. 
The 14\,m HERA dish design is a departure from the large N--small D approach used by 21\,cm 
observatories like PAPER and the MWA. Both observatories are actively pursuing power 
spectrum analyses using several year data runs, but the sheer data volume makes 
characterization and removal of systematics, as well as repeated or complimentary analyses, 
challenging. A larger element was chosen for HERA primarily to reduce correlation and data processing costs.

The HERA element is a 14\,m diameter parabolic dish with $f/D\approx0.32$ and dual-polarized dipole feed (Fig. \ref{fig:feedphoto}) suspended at prime focus, where $f/D$ is the ratio of focal length to diameter. The dish surface is formed by sheets of wire attached to PVC tubes attached to supports at the edge of the dish and to a hub at the base. The feed is suspended with cables attached to three telephone poles around the dish. The dipole stands 17'' below a 78'' diameter circular mesh back plane surrounded by a 30'' deep cylinder. The cylinder is offset 0.5'' from the back plane. The feed is designed to make the beams for both polarizations more similar to each other, and also to taper the dipole beam towards the edge of the dish to mitigate dish-dish coupling.  The nominal dish focus is at 4.5\,m from the dish surface, though numerical dish models including the feed predict the focus is closer to 5.3\,m.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Prototype HERA feed set up outside the dish for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17'' from a 78'' diameter back plane, surrounded by a 30'' deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

One cost of using fewer large antenna elements is a smaller field of view and thus fewer measured Fourier modes perpendicular to the line of sight. 
However, this is a 
small effect for 21\,cm power spectrum analyses as our leverage on $k$ modes comes primarily from modes along the line of sight (in the frequency dimension). Further, HERA's smaller field of view is actually desirable in that it drastically reduces the magnitude of emission at the edge of the wedge compared to a simple dipole element \citep{nithya15}. A second potential drawback is frequency structure introduced by time domain reflections between the suspended feed detailed by Ewall-Wice et al. (submitted) with simulations and Patra et al. (in prep) with zenith reflectometry measurements. These works demonstrate, though, that the slight frequency structure of the dish is small enough so as not to interfere with EOR science.

@@

We numerically model the HERA dish in both ANSYS HFSS\footnote{http://www.ansys.com/Products/Electronics/ANSYS-HFSS} and in CST Microwave Studio\footnote{https://www.cst.com/Products/CSTMWS}, making slightly different assumptions, in order to study the dependence of the beam pattern on modeling imperfections. In particular, we suspect that the near horizon beam response, which sets the level of horizon brightening, is quite sensitive to modeling assumptions. Modeling is also useful to probe the best focus of the dish--feed system. The nominal focus is at 4.5\,m given that the dish has $f/D=0.32$, though the feed back plane and skirt are expected to move the focus slightly.

The HFSS model (Fig. \ref{fig:modelbeams}, left panel) uses a 14\,m paraboloid with $f/D=0.32$ as the dish, with a 1\,m hole at the vertex. In reality the 1\,m diameter at the vertex is filled with concrete and supports the PVC pipes which suspend the mesh panels, but we effectively assume here it is perfectly absorbant.  The feed is modeled as a PAPER dipole suspended 17'' below a 78'' diameter circular backplane inside a 30'' deep skirt hanging 0.5'' below the backplane. All these surfaces are modeled as solid aluminum, as are the dipole ``sleeves'', circular disks just above and below the dipole plane designed to broaden its frequency response. The dipoles themselves is modeled as a copper pipe of diameter XX. In the simulations, we excite one of the dipoles using a modal port and measure the total gain response in each direction to 137.5\,MHz radiation.

In the CST model (Fig. \ref{fig:modelbeams}, middle panel) we again model the dish as a 14\,m diameter solid aluminum paraboloid with $f/D=0.32$ but now neglect the 1\,m hole at the vertex, effectively assuming the concrete and earth ground behind it are perfectly reflective. The actual material properties are somewhere in between, though difficult to predict theoretically. The feed model has the same dimensions as in the HFSS model. The simulations are done by exciting one of the dipoles with band-limited noise (100--200\,MHz). The beam pattern is then obtained at a given frequency using a farfield monitor after the excitation pulse energy within the structure decays to below -80\,dB. As expected, both models have slightly stronger sidelobes and wider main lobes than the ideal Airy pattern (Fig. \ref{fig:modelbeams}, right panel).

@@

\subsection{ORBCOMM Beam Mapping System Review}
\label{sec:orbcommreview}

We briefly review the beam mapping system detailed by \citet{neben15}, then discuss the 
application of system for HERA dish measurements. The system 
takes advantage of the 137\,MHz communications satellites operated by ORBCOMM Inc. 
as bright point sources which, by virtue of their number ($\sim30$), short orbital periods 
($\sim90$ minutes), and orbital precession, cover $\sim$65\% of the visible sky in just a few 
days. The coverage from the Green Bank site is limited by the fact that the satellites' orbital inclinations are all less 
than $45^\circ$.

@@

In detail, we measure the dual-polarization RMS power received by each antenna in 512 2\,kHz 
channels across the 137--138\,MHz band. Each band power is averaged over $\sim0.2$
\,sec. There are 0--3 satellites above the horizon at any given time transmitting on different 
$\sim15$\,kHz wide sub-bands in 137--138\,MHz. By observing at many different 
frequencies, we probe the beam response in all these directions simultaneously. We 
compute the satellite positions using the orbital elements published by Celestrak
\footnote{http://www.celestrak.com/NORAD/elements/orbcomm.txt} and the orbital 
integrator \texttt{predict}\footnote{http://www.qsl.net/kd2bd/predict.html}. However, the 
satellite frequencies vary occasionally to avoid interference within the constellation. 
\citet{zheng14} use interferometric phases to identify and exclude times when multiple 
satellites are in view. As our data acquisition system makes only total power 
measurements, we instead use an ORBCOMM interface box (typically supplied to 
commercial users of the network) to connect to passing satellites and record their identifier 
and transmission frequency during each pass.

In this way, beam measurements are built up along satellite tracks over the course of 
several days of integration, yielding typically 200--300 satellite pass. Each pass is 
processed separately to identify and exclude times of low signal-to-background when the 
satellite is low in the sky or in the off state of of a pulsing sequence. At those times, then 
satellite flux no longer dominates over that of the diffuse Galactic background, and a 
power measurement no longer probes the response in only the satellite direction. The beam 
measurements are then gridded in local Azimuth/Elevation coordinates in HEALPix \citep{healpix} as discussed in Sec. \ref{sec:orbcommreview}.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{ref_dipole_and_hera_dish.jpg}
\caption{The dish with its suspended feed is seen in the back, 50\,m north of one of the reference antennas used in the null experiment to study systematics. The experiment is conducted in Galford Meadow at NRAO--Green Bank.}
\label{fig:greenbankdishphoto}
\end{figure}

We use a simple dual-polarization dipole as our reference antenna. The dipole is constructed out of copper tubing covered by PVC for protection, mounted above a 2\,m $\times$ 2\,m ground plane. See \citet{neben15} for details. During the dish measurements the dipole is positioned 100\,m due south of the dish, though we experiment with other locations in order to characterize the environmental systematics of these measurements, as detailed in the next section. Figure \ref{fig:greenbankdishphoto} shows the dish with suspended feed 50\,m north of one of the reference antennas.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_maps.pdf}
\caption{Measured dish power patterns at various feed heights for the EW (left panel) and NS (right panel) linear instrumental polarizations. The sidelobes shrink and the main lobe narrows as the feed is raised, confirming that the focus of the dish/feed system is higher than the 4.5\,m ideal focus of the parabolic dish.}
\label{fig:measuredbeammaps}
\end{figure*}
